---
title: "Zillow_Case_01_Prelim-EDA_set2"
author: "Team ZillowCase"
date: "8/20/2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## EDA Visualization - Set II

This document assembles various visualizations that were helpful in preprocessing and preparations ahead of 
most of the real machine learning and model building for the data set.  Models in this file were presented as part
of EDA though some may get reused later in the research.

```{r setupCell1, warning=FALSE}
## libraries and other setup code may include things that were not actually used in this document
## it is amalgamated from many working files

#setwd('~/Desktop/zillow/input/')
library(corrplot)
library(ggplot2)
library(dplyr)
library(data.table)

library(ggplot2)
library(stringr)
library(DT)
library(tidyr)
library(lubridate)
library(stringi)
library(Hmisc)
library(car)

# train = fread('train_2016_v2.csv')
# properties = fread('properties_2016.csv')
# full = properties %>% left_join(., train, by = 'parcelid')

full_clean <- fread('clean_output.csv')  ## output from cleaning script: 2.3Million records (w and w/o logerror)

```


### Missingness and Data Exploration using Models And Other Visualizations

```{r}

## set up data ...
train_s1 = full_clean[!is.na(full_clean$logerror), ]
test_s1 = full_clean[is.na(full_clean$logerror), ]
# splitting out merged data frame into test and train
p = sample(1:nrow(train_s1), 7*nrow(train_s1)/10)
# making a sample
sub_train = train_s1[p, ]
# sub train is now the training portion of our train set
sub_test = train_s1[-p, ]
# sub test is now the test portion of our train set

```

#### EDA Model 1:

```{r}

model1 = lm(logerror ~ taxamount +
              calculatedbathnbr + 
              yearbuilt, data = sub_test)

summary(model1)  # look at summary for simple model

```

```{r}

plot(model1)
# a lot of violations in assumptions
```
Comments indicate many models were tried and only this one was kept:

```{r m2, eval=FALSE, echo=FALSE, include=FALSE}
## this model uses one or more fields from our data cleaning file.  It may have been created
## before full cleaning code was developed but is re-run here using the output of zillow_clean_<date>.R

model2 = lm(logerror ~ 
              calculatedfinishedsquarefeet +
              calculatedbathnbr + 
              airconditioningtypeid +
              buildingqualitytypeid +
              garagetotalsqft + 
              yearbuilt + 
              fireplacecnt + 
              unitcnt + 
              regionidcity +
              regionidzip + 
              regionidcounty +
              airconditioningtypeid +
              bedroomcnt + 
              lotsizesqft_imputed +
              poolcnt +
              decktypeid +
              numberofstories +
              basementsqft +
              yardbuildingsqft17 + 
              yardbuildingsqft26 +
              garagecarcnt + 
              taxdelinquencyflag +
              heatingorsystemtypeid +
              fullbathcnt +
              taxamount,
              data = sub_test)

# this is where all of my tinkering was done. A lot of models were created 
# here but I only included the final one 
summary(model2)

```

```{r}
## repeating same model to head off knitting issue
model2 = lm(logerror ~ 
              calculatedfinishedsquarefeet +
              calculatedbathnbr + 
              airconditioningtypeid +
              buildingqualitytypeid +
              garagetotalsqft + 
              yearbuilt + 
              fireplacecnt + 
              unitcnt + 
              regionidcity +
              regionidzip + 
              regionidcounty +
              airconditioningtypeid +
              bedroomcnt + 
              lotsizesqft_imputed +
              poolcnt +
              decktypeid +
              numberofstories +
              basementsqft +
              yardbuildingsqft17 + 
              yardbuildingsqft26 +
              garagecarcnt + 
              taxdelinquencyflag +
              heatingorsystemtypeid +
              fullbathcnt +
              taxamount,
              data = sub_test)
plot(model2)
# some issues here with the assumptions (QQ plot is s shaped)

```


```{r}
full_clean2 <- fread('clean_output_wDates.csv')  ## output from cleaning script: 2.3Million records (w and w/o logerror)
train_s2 = full_clean2[!is.na(full_clean2$logerror), ]

```
```{r tstcell2, eval=FALSE, echo=FALSE}
## quick test ... not output in markdown document
head(train_s2)
```


```{r}
## some experimenting done with this code ... only final experiment(s) shown

x = train_s2 %>% select(calculatedfinishedsquarefeet, 
                           calculatedbathnbr, 
                           airconditioningtypeid, 
                           buildingqualitytypeid, 
                           garagetotalsqft, 
                           yearbuilt, 
                           fireplacecnt, 
                           unitcnt, 
                           regionidcity, 
                           regionidzip, 
                           regionidcounty, 
                           airconditioningtypeid, 
                           bedroomcnt, 
                           lotsizesqft_imputed, 
                           poolcnt, 
                           decktypeid, 
                           numberofstories, 
                           basementsqft, 
                           yardbuildingsqft17,  
                           yardbuildingsqft26, 
                           garagecarcnt, 
                           taxdelinquencyflag, 
                           heatingorsystemtypeid, 
                           fullbathcnt, 
                           taxamount, 
                           # transactiondate,  # this field is not numeric
                           logerror)
x = x %>% select(-heatingorsystemtypeid, -fullbathcnt)
x$numberofstories = as.numeric(x$numberofstories)
x_whole = cor(x)
x_cor = corrplot(x_whole, method="shade",shade.col=NA, tl.col="black", tl.srt=45)

## code leading up to these plots
x_pos = x[x$logerror > 0, ]
x_neg = x[x$logerror < 0, ]
x2_pos <- cor(x_pos,  use = "pairwise.complete.obs")
x2_neg<- cor(x_neg,  use = "pairwise.complete.obs")

x_cor_pos = corrplot(x2_pos, method="shade",shade.col=NA, tl.col="black", tl.srt=45)
x_cor_neg = corrplot(x2_neg, method="shade",shade.col=NA, tl.col="black", tl.srt=45)
# correlations - positive and negative differ from each other so slightly it is easy to miss it

```

```{r}



```

```{r}


```


