---
title: "Team ZillowCase Missingness Exploration"
author: "Team ZillowCase"
date: "8/19/2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Missingness Non-Visual

This document assembles preliminary research about missingness within the Zillow data.  It contains field lists,
percents, and techniques for obtaining this information that do not include EDA Visualizations.  It was helpful in
decision making and in planning for visualizations in other documents


```{r setupCell1, warning=FALSE}
## libraries and other setup code may include things that were not actually used in this document
## it is amalgamated from many working files

#setwd('~/Desktop/zillow/input/')
library(corrplot)
library(ggplot2)
library(dplyr)
library(data.table)

train = fread('train_2016_v2.csv')
properties = fread('properties_2016.csv')
full = properties %>% left_join(., train, by = 'parcelid')
# sub=select(full, starts_with(c('finishedsquarefeet','logerror')))

```

```{r}
missing = lapply(full, function(i) sum(is.na(i))/nrow(full)) #to see missingness
missing$finishedfloor1squarefeet

```

```{r m1}

missing

```

```{r}
library(mice)
a=full[grepl('finishedsquare|logerror', names(full))]
md.pattern(a[-nrow(a)])
```

```{r}
full_set1 = full[!grepl('finishedsquare', 
                     names(full)[!grepl('calculatedfinishedsquare',names(full))])]

md.pattern(full_set1[grepl('garage', names(full_set1))]) #either none or both

```

```{r}
unique(full$hashottuborspa) #either blank or true
sum(full$hashottuborspa=="true") # 2% have hot 
```

```{r}
missing$taxvaluedollarcnt              #not too much missing
length(unique(full$taxvaluedollarcnt)) #continuous
unique(full$heatingorsystemtypeid)     #which values from data dictionary in use + NAs present

```


```{r}

full_set1$hashottuborspa[full_set1$hashottuborspa==""]="false"
full_set1$hasheat = !is.na(full_set1$heatingorsystemtypeid) #added column to say T/F for having heat based on NAs
md.pattern(full_set1[grepl('dollarcnt', names(full_set1))])

```

```{r}
full %>% select(regionidcity, latitude, longitude, fips) %>% md.pattern()
```
regioncityid is missing either when all 4 or missing or when the other 3 are present lattitude, longitude and fips always have values together so they can’t help w/ each other’s missingness

2922495 = rows where all 4 columns have values 51410 = rows where only regionidcity missing (from this set of 4 cols) 11437 = missing all 4 …

Final col: 0 cols missing data on this row 1 col missing data on this row 4 cols missing data on this row Total is totaling across rows subtotal so: 97158 data cells missing within this set of columns / rows (df)

```{r}
full %>% select(regionidcity, latitude, longitude, fips, 
                regionidcounty, regionidneighborhood, regionidzip) %>% md.pattern() # 

```

