---
title: "Team ZillowCase Preliminary EDA"
author: "Team ZillowCase"
date: "8/19/2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## EDA Visualization

This document assembles various visualizations that were helpful in preprocessing and preparations ahead of 
most of the real machine learning and model building for the data set.  

```{r setupCell1, warning=FALSE}
## libraries and other setup code may include things that were not actually used in this document
## it is amalgamated from many working files

#setwd('~/Desktop/zillow/input/')
library(corrplot)
library(ggplot2)
library(dplyr)
library(data.table)

library(ggplot2)
library(stringr)
library(DT)
library(tidyr)
library(lubridate)
library(stringi)
library(Hmisc)
library(car)

train = fread('train_2016_v2.csv')
properties = fread('properties_2016.csv')
full = properties %>% left_join(., train, by = 'parcelid')
# sub=select(full, starts_with(c('finishedsquarefeet','logerror'))) ## not working ...

```

### Missingness Related Corplots

```{r}
a=full[grepl('finishedsquare|logerror', names(full))]
corrplot(cor(a,use = "pairwise.complete.obs"))

```


```{r}

full_set1 = full[!grepl('finishedsquare', 
                     names(full)[!grepl('calculatedfinishedsquare',names(full))])]
full_set1$hashottuborspa[full_set1$hashottuborspa==""]="false"
full_set1$hasheat = !is.na(full_set1$heatingorsystemtypeid) #added column to say T/F for having heat based on NAs
corrplot(cor(full[grepl('dollarcnt', names(full))], use = "pairwise.complete.obs"))

```


```{r}

# help with using grep/grepl in this context:
#      https://awakeningdatascientist.wordpress.com/2015/07/20/r-of-the-day-grep-and-grepl/

tmpCols1 = full[grepl('regionidcity|latitude|longitude|fips', names(full))]
tmpCols1 = as.data.frame(sapply(tmpCols1, as.numeric))  # note: fips factor converted to level values: 1,2,3
tmpCols2 = full[grepl('region|fips|latitude|longitude', names(full))]
tmpCols2 = as.data.frame(sapply(tmpCols2, as.numeric))  # note: fips factor converted to level values: 1,2,3
## convert to numeric first to use in cor plot (see above):
corrplot(cor(tmpCols1, use = "pairwise.complete.obs"))

```

```{r}

corrplot(cor(tmpCols2, use="pairwise.complete.obs"))

```

 - Longitude shows a moderately strong negative correlation to latitude.
 - regionidcity shows a moderatly strong correlation to latitude
 - regionidcounty is showing a moderately strong correlation to Latitude
 - regionidcounty is showing a strong negative correlation to fips
 
 