---
title: "Zillow Case Project - Starter Code 1"
author: "TheMitchWorksPro"
date: "8/9/2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

```

## Code Testing for Data Cleaning
Set this up to test ideas for cleaning the code

```{r libSetup, warning=FALSE}
## library list from here:
# https://www.kaggle.com/philippsp/exploratory-analysis-zillow/notebook
library(data.table)
library(dplyr)
library(ggplot2)
library(stringr)
library(DT)
library(tidyr)
library(corrplot)
# library(leaflet)
library(lubridate)

## few more
library(stringi)

```

```{r}

# Import libraries
library(caret)

# Setting Parallel processing
library(doMC)
library(parallel)
number_of_cores <- detectCores()
registerDoMC(cores = number_of_cores/2)  ## further research:
    # http://blog.aicry.com/r-parallel-computing-in-5-minutes/

```


```{r}
### original values before adding skip blank rows argument for comparison ... none were eliminated
# Read 2985217 rows and 58 (of 58) columns from 0.604 GB file in 00:00:07
# [1] 2985217      58

datDir <- "./input/"
homePropertiesFile <- "properties_2016.csv"
logErrorDataFile <- "train_2016_v2.csv"

properties <- fread(paste0(datDir, homePropertiesFile), header=TRUE, na.strings=c("", " ", "NA", "N/A", "null"),
                   blank.lines.skip = TRUE,
                   colClasses=list(character=50))  ## eliminates warning so we know nothing else is wrong
                                                   ## na.string = common values for NA that may have been used
dim(properties)

train = fread(paste0(datDir, logErrorDataFile))
full = properties %>% left_join(., train, by = 'parcelid')



```

```{r}
head(train)

```


```{r}
### from NYC DSA Analysis:  correct Lat/Long to proper numbers
full <- full %>%
  mutate(latitude = latitude/1e6, longitude = longitude/1e6)
```


## Looking into Number of Stories

```{r}
missing = lapply(full, function(i) sum(is.na(i))/nrow(full)) # verity missingness
missing$numberofstories

```


```{r}
table(full$numberofstories)
unique(full$numberofstories)
```

```{r}
## this started as a look at the really strange properties with "too many stories"
## to determine if they were valid ...

## spot check suggests that properties with more than 5 are dirty data
## though logically you can have properties with more than 5 floors
## these 10 records do not have enough square feet or $ value to explain the extra floors

full %>% filter(as.integer(numberofstories) > 5) %>%                      
         select(parcelid, numberofstories, bathroomcnt,bedroomcnt,        
         buildingclasstypeid, calculatedfinishedsquarefeet, taxvaluedollarcnt, unitcnt)
   ## unitcnt is NA for all of these
                                          

```

By comparison, at 5 floors, some of the records look like they could support more, but without more information,
this data has to be treated as corect:

```{r}
full %>% filter(as.integer(numberofstories) == 5) %>%                      
         select(parcelid, numberofstories, bathroomcnt,bedroomcnt,        
         buildingclasstypeid, calculatedfinishedsquarefeet, taxvaluedollarcnt, unitcnt)
  ## at 5 floors, we see buildings with high $value and in some cases, unitcounts to corroborate they are probably apt. buildings.

```


```{r}

# stats_mode <- function(x) {
#   # pass in variable column and calculate the mode on it
#   names(which.max(table(train_data$x)))  ## assumes a global train_data
#                                          ## worked in Shu's code, broke in mine
# }

stats_mode <- function(x) {
  # pass in variable column and calculate the mode on it
  names(which.max(table(x)))
}

# stats_mode2 <- function(x) {
#   # pass in variable column and calculate the mode on it
#   names(which.max(table(full$x)))
# }
# full %>% stats_mode2(., numberofstories)

stats_mode(full$numberofstories)


```

```{r}

mode1 = stats_mode(full$numberofstories)

full <- full %>%  
  mutate(numberofstories = ifelse(is.na(numberofstories), mode1, numberofstories)
  )

table(full$numberofstories)  # check that table of values unchanged by this rule
unique(full$numberofstories) # check that NAs are now gone

```

```{r}
## separate rule so we can comment this out if adding more data later for which this rule is no longer true
## in 2016 data, see research file showing that properties with floors greater than 5 appear to be dirty data
# full$numberofstories[numberofstories > 5]

full <- full %>%  
  mutate(numberofstories = ifelse(as.integer(numberofstories) > 5, mode1, numberofstories)
  )

table(full$numberofstories)  ## check that data changed as expected ...
unique(full$numberofstories)

```

lotsize:

```{r}
### test ####
library(Hmisc)
set.seed(0)
full$lotsizesqft_imputed <- Hmisc::impute(full$lotsizesquarefeet, "random")
summary(full)

```
```{r}
# mean only changed from 22823 to 22923 (not that significant)
## check change in sd ...
sd(full$lotsizesquarefeet , na.rm=T)
sd(full$lotsizesqft_imputed)

(sd(full$lotsizesqft_imputed) -  sd(full$lotsizesquarefeet , na.rm=T)) / sd(full$lotsizesquarefeet , na.rm=T)
## 11% change in sd()
```


```{r}

full %>% filter(lotsizesquarefeet < calculatedfinishedsquarefeet) %>%                      
         select(parcelid, lotsizesquarefeet , calculatedfinishedsquarefeet, 
                finishedfloor1squarefeet,
                finishedsquarefeet6,
                finishedsquarefeet12,
                finishedsquarefeet13,
                finishedsquarefeet15,
                finishedsquarefeet50
                ) %>% nrow()     ## comment out from %>% to view records in Markdown doc
   ## unitcnt is NA for all of these


```

Finish up Tax Delinquincy and Fields from starterCode1

```{r}
## taxdelinquencyyear  (repeat from doc1)
class(full$taxdelinquencyyear)
unique(full$taxdelinquencyyear)

table(full$taxdelinquencyyear)
```

```{r}
full <- full %>% mutate(taxdelinquencyyear4d = ifelse(is.na(taxdelinquencyyear), 0,
                                     ifelse(taxdelinquencyyear < 17, taxdelinquencyyear+2000, taxdelinquencyyear+1900)))

unique(full$taxdelinquencyyear4d)
table(full$taxdelinquencyyear4d)

```

```{r}
## repeate earlier rules from startercode1 to use them here as well
full <- full %>%  
  mutate(taxdelinquencyflag = ifelse(is.na(taxdelinquencyflag), 0, ifelse(taxdelinquencyflag =="Y", 1, taxdelinquencyflag)),
         hashottuborspa = ifelse(is.na(hashottuborspa), 0, ifelse(hashottuborspa =="true", 1, hashottuborspa)),
         fireplaceflag = ifelse(is.na(fireplaceflag), 0, ifelse(fireplaceflag =="true", 1, fireplaceflag)),
         fireplacecnt = ifelse(is.na(fireplacecnt), 0, fireplacecnt)
  )

full <- full %>%  
  mutate(taxdelinquencyflag = ifelse(!is.na(taxdelinquencyyear) & taxdelinquencyflag==0, 1, taxdelinquencyflag),
         fireplaceflag = ifelse(fireplacecnt > 0, 1, fireplaceflag),
         fireplacecnt = ifelse(fireplaceflag==1 & fireplacecnt == 0, -1, fireplacecnt)  
         # -1 is a marker indicating that we need to impute these values if they exist
         # none were actually found when code was run but that may change with 2017 data in October
  )

unique(full$taxdelinquencyflag)

```


```{r}
## check done ... no rule needed yet
full %>% select(taxdelinquencyflag, taxdelinquencyyear, taxdelinquencyyear4d) %>% 
  filter(is.na(taxdelinquencyyear) & taxdelinquencyflag==1)  ## verified w/ other combination
                                                              ## nothing to return                         
```

```{r}
## check done ... no rule needed yet
full %>% select(taxdelinquencyflag, taxdelinquencyyear, taxdelinquencyyear4d) %>% 
  filter(!is.na(taxdelinquencyyear) & taxdelinquencyflag==0)  ## verified w/ other combination
                                                              ## nothing to return
```

```{r}
## quick check before moving on
unique(full$taxdelinquencyyear4d)
table(full$taxdelinquencyyear4d)

```

```{r}
## new field: calculate number of years delinquent:
## -----------------
## if 0, house is not in delinquency
## If 1 or more, house is delinquent and this is the x year of delinquincy
## example: 2016 will come out as 1 for first year even though first year is not complete
##          this one field should now reflect full delinquency state
##          update variable: curYear to 2017 after new data comes out

curYear = 2016

full <- full %>%  
  mutate(taxdelinquencyNoYrs = ifelse(taxdelinquencyyear4d == 0, 0, curYear - taxdelinquencyyear4d + 1))

```

```{r}
## spot check of values and ensure NAs are not present
unique(full$taxdelinquencyNoYrs)
table(full$taxdelinquencyNoYrs)

```

Fields from set 2 having to do with Location (and related fields to help with imputation)

```{r}
## rerun missing calculations
missing = lapply(full, function(i) sum(is.na(i))/nrow(full)) # verity missingness
full$fips <- as.factor(full$fips)  # make fips a factor
```


```{r}

"FIPS:"
levels(full$fips)
"FIPS Missing:"
missing$fips
"Latitude Missing:"
missing$latitude
"Longitude Records:"
missing$longitude

## from spotcheck of test in StarterCode1:
## check these for relationship
                          ## Missingness:
# 'regionidcounty'        ## 0.3831212  (same as Lat/Long/fips)
# 'regionidcity'          ## 2.1052071
# 'regionidzip'           ## 0.4683077
# 'regionidneighborhood'  ## 61.2623806

## function used forgot to multiply by 100 ...
## so 0.0038... below is same as 0.38% above (still less than 1% though)

```

```{r}
## quick test of assumptions:  
# how many records are missing all of these?: regionidcounty, latitude, longitude, fips? (miss % same)
library(mice)
full %>% select(regionidcity, latitude, longitude, fips) %>% md.pattern()

```

regioncityid is missing either when all 4 or missing or when the other 3 are present
lattitude, longitude and fips always have values together so they can't help w/ each other's missingness

2922495 = rows where all 4 columns have values
51410 = rows where only regionidcity missing (from this set of 4 cols)
11437 = missing all 4 ...

Final col: 0 cols missing data on this row
           1 col  missing data on this row
           4 cols missing data on this row
           Total is totaling across rows subtotal so: 97158 data cells missing within this set of columns / rows (df)

```{r}
# full %>% select(regionidcity, latitude, longitude, fips) %>% filter(is.na(regionidcity) & !is.na(latitude) 
#                                                                     & !is.na(longitude) & !is.na(fips)) 
#    ## above can help verify undersatnding of graph by switching is.na to !is.na ...

# library(mice)
full %>% select(regionidcity, latitude, longitude, fips,
                regionidcounty, regionidneighborhood, regionidzip) %>% md.pattern()

```

Are corplots of these for complete values only helpful?

```{r}
# help with using grep/grepl in this context:
#      https://awakeningdatascientist.wordpress.com/2015/07/20/r-of-the-day-grep-and-grepl/

tmpCols1 = full[grepl('regionidcity|latitude|longitude|fips', names(full))]
tmpCols1 = as.data.frame(sapply(tmpCols1, as.numeric))  # note: fips factor converted to level values: 1,2,3
tmpCols2 = full[grepl('region|fips|latitude|longitude', names(full))]
tmpCols2 = as.data.frame(sapply(tmpCols2, as.numeric))  # note: fips factor converted to level values: 1,2,3
## convert to numeric first to use in cor plot (see above):
corrplot(cor(tmpCols1, use = "pairwise.complete.obs"))
corrplot(cor(tmpCols2, use="pairwise.complete.obs"))

```
Analysis:

- Longitude shows a moderately strong negative correlation to latitude.
- regionidcity shows a moderatly strong correlation to latitude
- regionidcounty is showing a moderately strong correlation to Latitude
- regionidcounty is showing a strong negative correlation to fips

```{r}
### for later step in cleaning process ... basic code to convert all specified string columns to factors from the df

# full_2 <- full  # synch up my two copies before moving forward ...
# 
# factor_cols <- c("airconditioningtypeid", 
#                  "fips",
#                  "heatingorsystemtypeid", 
#                  "propertycountylandusecode",
#                  "propertylandusetypeid",
#                  "propertyzoningdesc", 
#                  "rawcensustractandblock",
#                  "regionidcity", 
#                  "regionidcounty", 
#                  "regionidneighborhood",
#                  "regionidzip",
#                  "censustractandblock")
# 
# ## convert columns to factors
# full_2 <- full_2 %>% 
#   mutate_at(.funs = as.factor, 
#             .vars = intersect(names(properties2), factor_cols))
```


```{r}


```
